\chapter{Related Projects}\label{c:related-work}

\section{My Health My Data}\label{s:mhmd}
My Health My Data project (MHMD) \cite{mhmd} is a Horizon 2020 Research and Innovation Action which aims at fundamentally changing the way sensitive data are shared.

MHMD is poised to be the first open biomedical information network centred on the connection between organizations and individuals, encouraging hospitals to start making anonymised data available for open research, while prompting citizens to become the ultimate owners and controllers of their health data.
It is an academic research project, focusing on the use of Privacy\hyp Preserving Data Mining (PPDM) tools in conjunction with Blockchain technologies, to develop an ecosystem where patient health records across several hospitals in Europe can be safely and traceably shared for the purposes of scientific research.
MHMD is intended to become a true information marketplace, based on new mechanisms of trust and direct, value-based relationships between EU citizens, hospitals, research centres and businesses.

The MHMD project currently utilizes the Sharemind secure computing platform, as well as many algorithms presented in the current thesis.



\section{Scalable Oblivious Data Analytics}\label{s:soda}
Scalable Oblivious Data Analytics project (SODA) \cite{soda} is also a Horizon 2020 Research and Innovation Action, with similar goals as MHMD project.
More specifically, the SODA project aims to tackle exactly data protection and anonymization issue identified by the Big Data Value Association (BDVA) \cite{bdva}, enabling practical privacy\hyp preserving analytics on Big Data by significant improvement of MPC techniques for privacy\hyp preserving Big Data processing.

One of the differences between the MHMD and the SODA project is that in the  latter uses FRESCO (a FRamework for Efficient and Secure COmputation) \cite{chan2017privacy, alxd1} as the underlying MPC engine.

